Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula

### 2024-02-15 output/paper/_si.qmd
Reviewing manuscript.

- Files for supplementary data must be accompanied by a summary of the file names and types.

@@

::: {#supptbl-lloq-mets}
```{r lloq-mets}
lloq_mets <- load_steroids()$loq |>
  tidylog::mutate(
    Metabolite = stringr::str_replace(
      metabolite,
      "^X", ""
    )
  ) |>
  tidylog::rename(
    LLOQ = "loq"
  ) |>
  tidylog::select(-metabolite, Metabolite, LLOQ)
gt::gt(lloq_mets |> dplyr::arrange(Metabolite)) |>
  gt::tab_footnote(
    footnote = "Abbreviations: lower limit of quantification (LLOQ)."
  ) |>
  gt::opt_footnote_marks(
    marks = "letters"
  )
```
**Lower limits of quantification expressed in ng/ml for the glucocorticosteroids (HELIX subcohort; 2013-2016).**
:::

@@

::: {#suppfig-flow-pop}
```{mermaid}
flowchart TB
  helixsc["HELIX subcohort\n(N = 1,301)"] --> edcs["HELIX data EDCs\n(N = 1,297)"]
  helixsc --> corts["HELIX data glucocorticosteroids\n(N = 1,004)"]
  edcs --> inter["HELIX data EDCs and glucocorticosteroids\n(N = 976)"]
  corts --> inter
  edcs -.-> rq1(["HRT-SE ~ EDCs"])
  inter -.-> rq2(["metabolites ~ EDCs"])
  inter -.-> rq3(["HRT-SE ~ metabolites"])
```
Flowchart describing thew sample size for each research question.
:::

@@

tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}}
 (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}.
:::

>>>>

@@

::: {#supptbl-lloq-mets}
```{r lloq-mets}
lloq_mets <- load_steroids()$loq |>
  tidylog::mutate(
    Metabolite = stringr::str_replace(
      metabolite,
      "^X", ""
    )
  ) |>
  tidylog::rename(
    LLOQ = "loq"
  ) |>
  tidylog::select(-metabolite) |>
  tidylog::select(Metabolite, LLOQ)
gt::gt(lloq_mets |> dplyr::arrange(Metabolite)) |>
  gt::tab_footnote(
    footnote = "Abbreviations: lower limit of quantification (LLOQ)."
  ) |>
  gt::opt_footnote_marks(
    marks = "letters"
  ) |>
  gt::tab_style(
    locations = gt::cells_column_labels(),
    style = list(gt::cell_text(weight = "bold"))
  )
```
**Lower limits of quantification expressed in ng/ml for the glucocorticosteroids (HELIX subcohort; 2013-2016).**
:::

@@

::: {#suppfig-flow-pop}
```{mermaid}
flowchart TB
  helixsc["HELIX subcohort\n(N = 1,301)"] --> edcs["HELIX data EDCs\n(N = 1,297)"]
  helixsc --> corts["HELIX data glucocorticosteroids\n(N = 1,004)"]
  edcs --> inter["HELIX data EDCs and glucocorticosteroids\n(N = 976)"]
  corts --> inter
  edcs -.-> rq1(["HRT-SE ~ EDCs"])
  inter -.-> rq2(["metabolites ~ EDCs"])
  inter -.-> rq3(["HRT-SE ~ metabolites"])
```
Flowchart describing the sample size for each research question.
:::

@@

tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  outcome = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```
**Marginal contrasts for effect modification by sex of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse >}} expressed in {{< acr mss >}}
 (HELIX subcohort; 2013-2016).** 
Circles and triangles indicate effect estimates. Solid lines indicate the $95\%$ {{< acr ci >}}. Abbreviations: cortisone production (cortisone prod.); cortisol production (cortisol prod.); corticost. prod. (corticosterone production).
:::

### 2024-02-15 output/paper/_tables.qmd
Reviewing manuscript.

gtsummary::tbl_merge(
  tbls = list(desc_sums_all, desc_sums_intersection)
) |>
  gtsummary::modify_spanning_header(
    dplyr::everything() ~ NA_character_
  ) |>
  gtsummary::as_gt() |>
  gt::tab_footnote(
    footnote = "Measurements available for the HELIX subcohort.",
    locations = gt::cells_column_labels(8)
  ) |>
  gt::opt_footnote_marks(
    marks = "letters"
  )
```
**Participants derived glucocorticosteroids concentrations expressed in ng/ml, by cohort and overall (HELIX subcohort; 2013-2016).**
:::

>>>>

gtsummary::tbl_merge(
  tbls = list(desc_sums_all, desc_sums_intersection)
) |>
  gtsummary::modify_spanning_header(
    dplyr::everything() ~ NA_character_
  ) |>
  gtsummary::as_gt() |>
  gt::tab_footnote(
    footnote = "Measurements available for the HELIX subcohort.",
    locations = gt::cells_column_labels(8)
  ) |>
  gt::opt_footnote_marks(
    marks = "letters"
  )
```
**Participants derived glucocorticosteroids concentrations expressed in ng/ml (HELIX subcohort; 2013-2016).**
:::

### 2024-02-15 output/paper/manuscript.qmd
Reviewing manuscript.

is_hpc <- FALSE
store <- ifelse(
  is_hpc, 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)
id_var <- "HelixID"
num_digits_est = 2
num_digits_sig = 2
params_ <- params(is_hpc = is_hpc)
threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params_$steps$rq2$preproc_outcome$missings$k
```

@@

Epidemiological evidence suggests that exposure to {{< acr edcs >}} may be linked to perturbation of the {{< acr hpa >}} axis, which has a major role in brain development. We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children. We used data from the {{< acr helix >}} cohort including 1,297 children aged 6-11. We used the parametric g-formula and {{< acr mcs >}} to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex. We observed positive {{< acr mcs >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}. Several {{< acr edcs >}} were also associated with positive {{< acr mcs >}} of cortisone, cortisol, and corticosterone production. Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found possible effect modification by sex for cortisone and corticosterone production. Our results suggest that multiple {{< acr edcs >}} might interfere with inattentiveness outcomes in children and with the homeostasis of the {{< acr hpa >}} axis. Further research is necessary to determine whether this can have a clinically significant impact on childhood neurodevelopment.

@@

- Authorship must follow the ICMJE's criteria. Author names should be listed in ScholarOne and author contributions should be detailed in the cover letter ("Author A..."). Neither names nor contributions should appear in the blinded manuscript. Do not include conflicts of interest.
- The length of the abstract must be less or equal than 200 words, and should be unstructured, stating the research questions, the methods used, and the results and conclusions of the research.
- The length must be <4,000 words, excluding abstract, references, tables, figure captions, acknowledgments, and Supplementary Material.
- Submission: SER members receive a 10% discount on the fees per page. This should be requested in the cover letter. Also 20% discount for Open Access charges. If uploading a single file (document, tables, figures, SM), designate the file as "Main document - anonymous".

The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants may play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. {{< acr edcs first_use=true >}}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023]. Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to {{< acr edcs >}} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].

@@

Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes, might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
In the present study, we thus estimated associations between 1) non-persistent {{< acr edcs >}} and attention, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large cohort in Europe.

@@

For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a {{< acr dag >}} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023]. The sets of covariates were selected to estimate the total effect of the exposure on the outcome. For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects. Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure. The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty. Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].

- For RQ1 I used creatinine values from HELIX. For RQ3 the ones from the steroids dataset. For RQ2, I included in the model both variables.

@@

Concentrations of the glucocorticosteroids were classified as quantifiable, below the {{< acr loq >}}, possible interference or out of range, and not detected. For each metabolite, we computed the fraction of values below the {{< acr loq >}} and not detected, both within each cohort and overall. We proceeded to impute these values using half the value of the corresponding {{< acr lloq >}}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall. Information about the {{< acr lloq >}} for the glucocorticosteroids is provided in @supptbl-lloq-mets. The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall. We used `r knn_k` nearest neighbors. We natural log-transformed them to improve model fit, assessed with posterior predictive checks. To do so, replicated data were simulated with the fitted models and compared to the observed data. We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021]. Values of cortisol production and cortisone production were expressed in nanograms per millilitre (ng/ml), whereas values of cortisol metabolism and 11bHSD activity were unitless.

@@

Missing values in the covariates were imputed similarly using kNN. Categorical covariates were imputed using the `maxCat` function, which chooses the level with the most occurrences. Creatinine values were expressed in grams per litre.

@@

We fit the outcome model using the `glm` function and a Gaussian family with identity link from base R. The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the {{< acr adrf >}}. When the outcome was a ratio, as was the case for cortisol metabolism and 11bHSD activity, we included the logarithm of its denominator, cortisol and cortisol production, respectively, as a control variable [@BartlettPartnoy:2020].

@@

We further estimated the {{< acr adrf >}} using the `avg_predictions` function from the `marginaleffects` R package, examining 50 exposure values from the 10th to the 90th percentiles of the exposure. As done for the {{< acr mcs >}}, we included the estimated balancing weights and used cohort as a clustering variable when computing robust standard errors.

@@

@Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}. For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}}. Most of the {{< acr cis first_use=true >}} included the null effect, though. Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`). The {{< acr opp >}} DETP was strongly negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).

@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on cortisone production, cortisol production, and corticosterone production. For most {{< acr edcs >}}, a cohort-specific increase in the levels of the exposures from the 10th to the 90th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites. Exceptions were BUPA, which was associated with negative {{< acr mcs >}} for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
The majority of the effects for the phenols and phthalate metabolites included the null. The phenol BPA showed the largest {{< acr mcs >}} across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects2, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects2, "bpa", "corticosterone production")`).

@@

@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure. Significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`). The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-1sa.

@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure and outcome. Significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes. The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)). The forest plots of the individual {{< acr mcs >}} are presented in @suppfig-marginal-2sa.

@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for each exposure. Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost. prod.", "HRT-SE")`). Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females). The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-3sa.

>>>>

is_hpc <- FALSE
store <- ifelse(
  is_hpc, 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)
id_var <- "HelixID"
num_digits_est = 3
num_digits_sig = 2
params_ <- params(is_hpc = is_hpc)
threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params_$steps$rq2$preproc_outcome$missings$k
```

@@

Epidemiological evidence suggests that {{< acr edcs >}} may perturb the {{< acr hpa >}} axis, which has a major role in brain development. We aimed to evaluate the effects of childhood exposure to organophosphate pesticides, phenols, and phthalate metabolites, on urinary glucocorticosteroids and inattention measures in children using data from the {{< acr helix >}} cohort. We used the parametric g-formula to estimate effects between {{< acr edcs >}}, glucocorticosteroids, and {{< acr hitrtse >}}, a measure of efficiency from the {{< acr ant >}}, and tested for possible effect modification by sex. We observed a positive {{< acr mc >}} for exposure increases from the 10th to the 90th percentile for methyl-paraben (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`), on {{< acr hitrtse >}}. Several {{< acr edcs >}} were also associated with a positive {{< acr mc >}} for cortisone, cortisol, and corticosterone production. Increased levels of the glucocorticosteroids had no effect on {{< acr hitrtse >}}, although we found a possible effect modification by sex. Our results suggest that multiple {{< acr edcs >}} might interfere with inattentiveness outcomes in children and with the homeostasis of the {{< acr hpa >}} axis.

@@

The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants may play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. Multiple {{< acr edcs first_use=true >}}, ubiquitous chemicals present in many every-day products and diet, are capable of interfering with the endocrine system, and have shown associations with childhood neurodevelopment and behavior [@BouchardBellingerWright:2010; @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015; @HuangChenSu:2015; @CartierWarembourgLeManer-Idrissi:2016; @TewarAuingerBraun:2016; @YuDuChiou:2016; @HuangTsaiChen:2017; @KimHongShin:2017; @FurlongHerringBuckley:2017; @Braun:2017; @LiZhangKuang:2018; @Rodriguez-CarrilloMustielesPerez-Lobato:2019; @ShoaffCoullWeuve:2020; @OhKimKannan:2023; @VilmandBeckBilenberg:2023]. Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to {{< acr edcs >}} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].

@@

Taken together, these results suggest that the negative influence of exposure to certain {{< acr edcs >}} on neurodevelopmental outcomes might be mediated, at least partially, by disruption of the {{< acr hpa >}} axis' homeostasis.
In the present study, we thus estimated associations between 1) non-persistent {{< acr edcs >}} and attention, 2) non-persistent {{< acr edcs >}} and glucocorticosteroids, and 3) glucocorticosteroids and attention, using the parametric g-formula and {{< acr mcs first_use=true >}}, in children of a large cohort in Europe.

@@

For each research question, defined by a specific type of exposure and outcome, the minimal set of covariates for inclusion in the analyses was selected on the basis of a {{< acr dag >}} built with `DAGitty` [@TextorvanderZanderGilthorpe:2016] and `ggdag` [@Barrett:2023]. The sets of covariates were selected to estimate the total effect of the exposure on the outcome. For effect estimation of the {{< acr edcs >}} on glucocorticosteroids and of glucocorticosteroids on {{< acr hitrtse >}}, these sets were also sufficient to estimate direct effects. Sample-specific creatinine values were used to adjust for possible dilution effects. Further, each minimal adjustment set was *augmented* with precision covariates, defined as the set of parents variable of the outcome that are not parents of the exposure. The adjustment sets are provided in the Supplementary Material as text files compatible with DAGitty. Codebooks for the used covariates, by research question, are provided in Supplementary Tables [-@supptbl-codebook-1; -@supptbl-codebook-2; -@supptbl-codebook-3].

@@

Concentrations of the glucocorticosteroids were classified as quantifiable, below the {{< acr loq >}}, possible interference or out of range, and not detected. For each metabolite, we computed the fraction of values below the {{< acr loq >}} and not detected, both within each cohort and overall. We proceeded to impute these values using half the value of the corresponding {{< acr lloq >}}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall. Information about the {{< acr lloq >}} for the glucocorticosteroids is provided in @supptbl-lloq-mets. The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall. We used `r knn_k` nearest neighbors. We natural log-transformed them to improve model fit, assessed with posterior predictive checks. To do so, replicated data were simulated with the fitted models and compared to the observed data. We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021]. Values of total cortisol, cortisone, and corticosterone production were expressed in nanograms per millilitre (ng/ml).

@@

Missing values in the covariates were imputed similarly using kNN. Categorical covariates were imputed using the `maxCat` function, which chooses the level with the most occurrences. Creatinine values were expressed in grams per litre (g/L).

@@

We fit the outcome model using the `glm` function and a Gaussian family with identity link from base R. The exposure variable was modeled using natural cubic splines with 3 degrees of freedom, to more flexibly capture the {{< acr adrf >}}.

@@

@@

@Fig-marginal-1 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on {{< acr hitrtse >}}. For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the values of {{< acr hitrtse >}}. Most of the {{< acr cis first_use=true >}} included the null effect, though. Statistically significant effects were observed for the phenol MEPA (`r format_res(meffects1, "mepa", "HRT-SE")`), and the phthalate metabolites oxo-MiNP (`r format_res(meffects1, "oxominp", "HRT-SE")`), oh-MiNP (`r format_res(meffects1, "ohminp", "HRT-SE")`), and MEHP (`r format_res(meffects1, "mehp", "HRT-SE")`). The {{< acr opp >}} DETP was negatively associated with {{< acr hitrtse >}} (`r format_res(meffects1, "detp", "HRT-SE")`).

@Fig-marginal-2 presents the forest plot for the {{< acr mcs >}} of the non-persistent {{< acr edcs >}} on total cortisone, cortisol, and corticosterone production. For most {{< acr edcs >}}, a cohort-specific decrease in the levels of the exposures from the 90th to the 10th percentiles was associated with a positive {{< acr mc >}}, indicating an increase in the total production of these metabolites. Exceptions were BUPA, which was associated with negative {{< acr mcs >}} for all three outcomes, and MiBP, which was associated with a negative {{< acr mc >}} for total cortisone production only.
The majority of the effects for the phenols and phthalate metabolites included the null. The phenol BPA showed the largest {{< acr mcs >}} across all three outcomes (cortisone production, `r format_res(meffects2, "bpa", "cortisone production")`; cortisol production, `r format_res(meffects2, "bpa", "cortisol production")`; corticosterone production, `r format_res(meffects2, "bpa", "corticosterone production")`).

@@

@Tbl-hypothesis-1and2 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for the {{< acr edcs >}} on the glucocorticosteroids and {{< acr hitrtse >}}. For {{< acr hitrtse >}}, significant differences were present for the phenol OXBE (`r format_res(heffects1, "oxbe", "HRT-SE")`) and the phthalate metabolites MEP (`r format_res(heffects1, "mep", "HRT-SE")`) and MbZP (`r format_res(heffects1, "mbzp", "HRT-SE")`). For the glucocorticosteroids, significant differences were present across all three classes of {{< acr edcs >}} and for all outcomes. The largest differences were attributable to the {{< acr opps >}} DMTP (cortisol production, `r format_res(heffects2, "dmtp", "cortisol production")`) and DETP (corticosterone production, (`r format_res(heffects2, "detp", "corticosterone production")`); cortisone production, (`r format_res(heffects2, "detp", "cortisone production")`)). The forest plots of the individual {{< acr mcs >}} are presented in Supplementary Figures -@suppfig-marginal-1sa and -@suppfig-marginal-2sa.

@Tbl-hypothesis-3 presents the results of a hypothesis test for the difference between estimates of the {{< acr mcs >}} for females and males, for the glucocorticosteroids on {{< acr hitrtse >}}. Significant differences were present for cortisone production (`r format_res(heffects3, "cortisone prod.", "HRT-SE")`) and corticosterone production (`r format_res(heffects3, "corticost. prod.", "HRT-SE")`). Furthermore, for all exposures, the {{< acr mcs >}} had opposite sign (positive for males and negative for females). The forest plot of the individual {{< acr mcs >}} is presented in @suppfig-marginal-3sa.
